\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

In this report, I present an integration of a road generation algorithm, described in \cite{roadgen}, with the GPU accellerated snow simulator developed at the HPC lab at the Department of Computer and Information Science at the Norwegian University of Science and Technology (NTNU). I also present a USGS DEM to RAW heightmap converter, for using real-world maps in the snow simulator. 

Given a height map of a terrain, my road generation application generates a road that minimizes a cost function that depends on the length of the road, slope and curvature. This is accomplished using an optimized A* search, generating a discrete shortest path represented by a series of nodes. 

The road trajectory is computed by using the nodes of the discrete shortest path as control points in a clothoid spline, which gives pleasant curves in the road. The trajectory is used to generate a RoadXML file, which is a description of the road (trajectory, road profile, textures, etc.); this is then used in SCANeR Studio developed by Octal\cite{octalstudio} to generate a triangle mesh of the road. This mesh is loaded into the snow simulator for rendering. The terrain is adjusted to accommodate the road model as neccesary, by constructing a distance map for each vertex in the road mesh, then adjusting the terrain elevation according to the closest vertices.

I present results showing visually pleasing rendering of the roads, that looks natural in the terrain. I also present performance data where different grid resolutions are tested; these results show that for many terrains, a lower resolution may not give a much higher cost of the trajectory, while giving immensely faster searches. In addition to this, I present performance results on using Dijkstra's algorithm vs. an A* search for generating the roads. The results show a clear difference, A* being up to 2.3 times faster, averaging at 1.7 times faster using the four test maps.

