\chapter{Introduction}
\label{chap:introduction}
In this report, I present my work of integrating procedurally generated roads with a snow simulator developed at the High Performance Computing lab at NTNU. I implement the road generation algorithm from \cite{roadgen}, output a RoadXML file\cite{roadxml}, use an external tool to generate a road mesh, and finally import this mesh into the snow simulator. 

The first part of this project, generating the road trajectory, is done using an A* graph search with points on a height map as the nodes, i.e. we perform a discrete shortest path search over the (potentially) continuous domain that is the terrain. Then, in order to generate a RoadXML file, the actual trajectory is computed from the discrete shortest path by using the nodes along the path as control points for a clothoid spline. The clothoid spline is constructed using the method described in section \ref{sec:back_clothoid}.

A road mesh is generated using SCANeR Studio from Octal, by importing the RoadXML file. A triangle mesh is output in the Wavefront OBJ format, which is then parsed by the snow simulator. Finally, the terrain is adjusted so it will match the road model; this is necessary due to the fact that the road trajectory elevation is sampled at discrete intervals when generating the RoadXML file, and that the road has a width which is not accounted for in the RoadXML.

\section{Project scope}
\label{sec:project-scope} 
The scope of this project is first and foremost to implement the road generator in \cite{roadgen}, and load these roads as meshes into the snow simulator. The project aims to test the performance of the road generation algorithm at different grid resolutions as well as comparing it with Dijkstra's algorithm, and present visual results showing roads rendered in the snow simulator. It is beyond the scope of this project to fully integrate the road generation algorithm, as proposed in chapter \ref{chap:futurework}, Future Work. Instead, an application external to the snow simulator is developed that generates roads and outputs a RoadXML file.

\section{Report organization}
First, the report will go through some of the background material used in this project in chapter \ref{chap:background}; this includes solving shortest path problems, heuristic search, an overview on the NVIDIA GPU architectures and road generation. Then, the implementation of the road generator (which includes the A* search), integration with the snow simulator and the USGS DEM parser is presented in chapter \ref{chap:implementation}. Visual results as well as performance results are presented in chapter \ref{chap:results}. Finally, chapter \ref{chap:futurework} summarizes some of the possible improvements and future projects, and chapter \ref{chap:conclusion} concludes.
